---
title: "5. Differential expression analysis"
author: "Yang Li"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(RColorBrewer)
library(Polychrome)
library(ggplot2) 
#library(tidyverse)
#library(enrichR)
#library(GSVA)
#library(ggpubr)
#library(Hmisc)
#library(ggpubr)
#library(ggrepel)
#library(ggfortify)
#library(QUBIC)
#library(DESeq2)
#library(pheatmap)
#library(patchwork)
#library(monocle)
#library(DelayedArray)
#library(DelayedMatrixStats)
#library(org.Mm.eg.db)
#library(VennDiagram)

#library(garnett)
#knitr::opts_knit$set(root.dir = "/fs/project/PCON0005/cankun/faith")
#
knitr::opts_knit$set(root.dir = "C:/Users/flyku/Desktop/pop")
source("functions.R")

```

```{r, echo = T, eval = T, message = F, warning = F, error = F}
#### 

#the faith_scrnaseq.RData contains all running results, load this data to save time
load("faith_scrnaseq_0602.RData")
#load("faith_scrnaseq.RData")

####

```

# Overview

Analyses:

Task : Find differentially expressed genes for the following comparisons for three cell types
(monocytes, astrocytes and oligodendrocyte precursors). Please show the cut-offs used
for parameter thresholds and types of statistical tests used. Please share the output
data tables to Faith’s shared folder.


```



Find DE genes for selected groups.

## Monocytes (cluster 5)

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

DefaultAssay(combined) <- 'RNA'
this_ct <- "5" # the cluster ID
this_obj <- subset(combined, subset = seurat_clusters == this_ct) # extract this cluster
Idents(this_obj) <- this_obj$orig.ident
levels(as.factor(this_obj$orig.ident))

```

### Sham Vehicle vs Sham PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + Sham"
group2 <- "PLX + ShamI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct,"_",group1,"_vs_",group2,".csv")
write.csv(this_markers, output_filename)

```

### 7d SCI Vehicle vs 7d SCI PLX5622

```{r, fig.width = 10, fig.height = 10,echo = T, message = T, warning = F}

group1 <- "Vehicle + 7d SCI"
group2 <- "PLX + 7d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct,"_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 14d SCI Vehicle vs 14d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 14d SCI"
group2 <- "PLX + 14d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct,"_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 28d SCI Vehicle vs 28d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 28d SCI"
group2 <- "PLX + 28d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```


## Astrocytes (cluster 6)

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}
this_ct <- "6"
this_obj <- subset(combined, subset = seurat_clusters == this_ct)
Idents(this_obj) <- this_obj$orig.ident
levels(as.factor(this_obj$orig.ident))

```

### Sham Vehicle vs Sham PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + Sham"
group2 <- "PLX + ShamI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1,"_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 7d SCI Vehicle vs 7d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 7d SCI"
group2 <- "PLX + 7d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1, "_vs_", group2, ".csv")
write.csv(this_markers,output_filename)

```

### 14d SCI Vehicle vs 14d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 14d SCI"
group2 <- "PLX + 14d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 28d SCI Vehicle vs 28d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 28d SCI"
group2 <- "PLX + 28d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

## Oligodendrocyte precursors (cluster 14)

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

this_ct <- "14"
this_obj <- subset(combined, subset = seurat_clusters == this_ct)
Idents(this_obj) <- this_obj$orig.ident
levels(as.factor(this_obj$orig.ident))

```

### Sham Vehicle vs Sham PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + Sham"
group2 <- "PLX + ShamI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 7d SCI Vehicle vs 7d SCI PLX5622

```{r, fig.width = 10, fig.height = 10,echo = T, message = T, warning = F}

group1 <- "Vehicle + 7d SCI"
group2 <- "PLX + 7d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct,"_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 14d SCI Vehicle vs 14d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 14d SCI"
group2 <- "PLX + 14d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct,"_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```

### 28d SCI Vehicle vs 28d SCI PLX5622

```{r, fig.width = 10, fig.height = 10, echo = T, message = T, warning = F}

group1 <- "Vehicle + 28d SCI"
group2 <- "PLX + 28d SCI"
this_markers <- Seurat::FindMarkers(this_obj, ident.1 = group1, ident.2 = group2)
output_filename <- paste0("Cluster", this_ct, "_", group1, "_vs_", group2, ".csv")
write.csv(this_markers, output_filename)

```



# Session Infomation

```{r}

sessionInfo()
# Print version information about R, the OS and attached or loaded packages.

```